---
title: "Chapter 4 Assignment"
author: "Jacob Smith"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Question 1
```{r}
library(tidyverse)




df <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/nfl_salaries.csv")
```
I just imported the data here, it was light work.

##Question 2
```{r}
df_tidy <- df %>% 
  pivot_longer(names_to = "position", 
               values_to = "salary", 
               cols = -year)
```

As the data itself was pretty ugly, i tided it up with pivot_longer and put all data within 3 columns

#Question 3
```{r}
quarterbacks <- df_tidy %>% 
  filter(position %in% "Quarterback")

ggplot(data = quarterbacks, mapping = aes(x = salary)) +
  geom_histogram(bins = 35) +
  facet_wrap(~ year)

```
The most evident pattern was the fact that the large majority of salaries were in the lower end, while there were a fewer on the higher end.

##Question 4

```{r}
df_tidy <- drop_na(df_tidy)
average_pay <- df_tidy |> 
  group_by(position, year) |> 
    summarize(mean = mean(salary))
```

To provide clarity, I then grouped the tidy dataframe by position and year. Those were then averaged to provide a picture into the pay of football players.


#Question 5     
```{r}
ggplot(average_pay, aes(x = year, y = mean, col = position)) +
  geom_line(position = "dodge") +
  labs (x = "Year", y = "Average Salary")
```

These were the mapped to provide a clearer image across different positions. The two immediate observations were that the valuations of both defense and offensive line-man increased at relatively similar rates starting in 2016. Secondly, the valuation of special teams, tight ends, and running backs have been relatively stagnant. Running backs even have decreased in value over the years. Overall, these track with the ongoing optimization of salaries due to analytics taking priority in salary management