---
title: "Chapter 1 Assignment"
output:
  html_document:
    df_print: paged
date: "2022-09-04"
---


```{r}
library(causact)
library(dplyr)
library(igraph)
```
##Question 3

```{r}
```

As the assignment asked, I ran "as_data_frame", using the parameters of the dplyr version. It gave me an error because as the message from the earlier section indicated, igraph masked "dpylr." Naturally, I have to run it as the dpylr version. 

```{r}
df <- dplyr::as_data_frame(x = c(1,2,3))
glimpse(df)
```
##Question 3

```{r}
x = c(5,6,2,7,9,1)
n_distinct(x)
```
Question 3 asked me to use and define n_distinct, which pretty simply returns the amount of unique values within a vector. When applied to "x", n_distinct returns 6 because are 6 unique valuees.

##Question 4
```{r}
glimpse(baseballData)
bb = baseballData
```
Running glimpse on the data reveals that there are 5 columns and 12,145.Those 5 columns refer to $Date, $Home, $Visitor, $HomeScore, $VisitorScore. Home refers to the home team. Homescore refers to the score of the home team in that game.
##Question 5
```{r}
baseballData[1,]
```
In this instance, one row in the data represents an individual game.

```{r}
baseballData[,2:3] %>% head()

```

If we wanted to select certain columns, such as $Home and $Visitor, then we can select them by using the space after the comma. These two columns represent the two teams in each individual game, the home team and the away team.

##Question 6
```{r}
name <-
  c(
    "Wayne Gretzky",
    "Gordie Howe",
    "Jaromir Jagr",
    "Brett Hull",
    "Marcel Dionne",
    "Phil Esposito" ,
    "Mike Gartner",
    "Alex Ovechkin",
    "Mark Messier" ,
    "Steve Yzerman"
  )

goals <- c(894, 801, 766, 741, 731, 717, 708, 700, 694, 692)

year_started <- c(1979, 1946, 1990, 1986, 1971, 1963, 1979, 2005, 1979, 1983)
df1 <- tibble( 
  player_names = name, 
  goals_scored = goals,
  yearstarted = year_started) 
glimpse(df1)
```
After creating a new data frame, creatively called df1, I successfully glimpsed the data frame created via tibble. It included what I expected, that is the information given to me.
