---
title: "SR CH3 Assignment"
output: html_document
date: "2022-09-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
devtools::install_github("rmcelreath/rethinking")

```

```{r}
library(rethinking)
library(tidyverse)
```


```{r}
n <- 1000
n_success <- 6
n_trials  <- 9

(
  d <-
  tibble(p_grid = seq(from = 0, to = 1, length.out = n),
         # note we're still using a flat uniform prior
         prior  = 1) %>% 
  mutate(likelihood = dbinom(n_success, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior = (likelihood * prior) / sum(likelihood * prior))
)  
  # how many samples would you like?
n_samples <- 1e4

# make it reproducible
set.seed(3)

samples <-
  d %>% 
  slice_sample(n = n_samples, weight_by = posterior, replace = T)

d %>% 
  slice_sample(n = 1e6, weight_by = posterior, replace = T) %>% 
  ggplot(aes(x = p_grid)) +
  geom_density(fill = "black") +
  scale_x_continuous("proportion of water (p)", limits = c(0, 1))
```
set up
### Easy Questions

## Question 1
```{r}
d %>% 
  filter(p_grid < .2) %>% 
  summarise(sum = n() / n_samples)

```

Only 2% of it lies below .02

## Question 2

```{r}
samples %>%
  filter(p_grid > .8) %>% 
  summarise(sum = n() / n_samples)
```

Only 12.4% lies above .8

## Question 3

```{r}
samples %>%
  filter(p_grid > .2 & p_grid < .8) %>%  
  summarise(sum = n() / n_samples)
```

As basic math would indicate, the other 87.6 percent fits within those two points.

## Question 4
```{r}
(q_20 <- quantile(samples$p_grid, prob = .2))


```

20% of the posterior probability lies abolive .5215215.


## Question 4

```{r}
(q_80 <- quantile(samples$p_grid, prob = .8))

```
Inversely, 20% of posterior probability lies above the p value of .7627628


## Question 6

```{r}

rethinking::HPDI(samples$p_grid, prob = .66)

```
As requested, the narrowest interval of 66% is from  0.5255255 to 0.7947948 

## Question 7

```{r}
library(tidybayes)

median_qi(samples$p_grid, .width = .66)

```
The values of P for an 66% interval that is equal before and after, is 0.5045045 to 0.7757758

### Medium Questions

## Question 1

```{r}
n_success <- 8
n_trials  <- 15

new_d <-
  d %>% 
  mutate(likelihood = dbinom(n_success, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior = (likelihood * prior) / sum(posterior))
set.seed(3)




new_d %>% 
ggplot(aes(x = p_grid, y = posterior)) +
  geom_point() +
  geom_line() +
  labs(x = "probability of water",
       y = "posterior probability") +
  theme()
```
This just changes the code done above. It produces a slightly different output.

## Question 2

```{r}


new_samples <-
  new_d %>% 
  slice_sample(n = n_samples, weight_by = posterior, replace = T)

HPDI(new_samples$p_grid, prob = .9)
```
The 90 percent HPDI is betweeen .3393393 and 0.7267267

## Question 3

```{r}
test <- rbinom(10000, size = 15, prob = new_samples$p_grid)
mean(test == 8)
simplehist(test)
       
```
14.53% chance for it to be 8 observations of water
## Question 4

```{r}
test2 <- rbinom(10000, size = 9, prob = new_samples$p_grid)
mean(test2 == 6)
simplehist(test2)
```

This one follows the same logic. It would result in a probability of 17.63 %. 

## Question 6
Time to redo them all.......... It makes me sad

The set up and 3M1
```{r}


(
  d8 <-
  tibble(p_grid = seq(from = 0, to = 1, length.out = n),
         prior  = ifelse(p_grid< .5, 0, 1)) %>% 
  mutate(likelihood = dbinom(n_success, size = n_trials, prob = p_grid)) %>% 
  mutate(posterior = (likelihood * prior) / sum(likelihood * prior))
)  

newnewsamples  <-
  d8 %>% 
  slice_sample(n = n_samples, weight_by = posterior, replace = T)


d8 %>% 
ggplot(aes(x = p_grid, y = posterior)) +
  geom_point() +
  geom_line() +
  labs(x = "probability of water",
       y = "posterior probability") +
  theme()
```
# Redo 3M2
```{r}
HPDI(newnewsamples$p_grid, prob = 0.9)
```
The new HPDI goes from 0.5005005 to 0.7127127, an obviously narrowwer section.

# Redo 3M3

```{r}
test3 <- rbinom(10000, size = 15, prob = newnewsamples$p_grid)
mean(test3 == 8)
simplehist(test3)
```

This one seemed to slightly shift the direction slightly to the right. It for example, now included 15 as an option. Furthermore, the percentage went to 15.69 percent which was a slight increase.

# Redo 3M4
```{r}
test4 <- rbinom(10000, size = 9, prob = newnewsamples$p_grid)
mean(test4 == 6)
simplehist(test4)
```
This was one resulted in a larger increase. It went from 17.47% to 23.26%.


## 3M6

```{r}

#????

```
Utterly unsure of how to do this one.